### 2018-03-15 submission/manuscript.Rmd
change fig legend order

**Figure 4: Comparison of Random Forest modeling approaches to classify individuals as having normal colons or adenomas (A) or carcinomas (B) when training the models using the taxa with significant ORs, all taxa in a community, or all OTUs in a community when using stool samples.** No taxa had a significant OR associated with the presence of adenomas using stool samples. The black line represents the median AUC for the respective group. The dashed gray line indicates an AUC of 0.5.

**Figure 5: Relative importance of taxa with significant ORs in Random Forest models for differentiating between individuals with normal colons and carcinomas using stool samples (A) or unmatched tissue samples (B).** The colors indicate the z-transformed (i.e. mean of 0.0 and standard deviation of 1.0) mean decrease in accuracy values calculated from the model for each study. The taxa are ranked by their mean z-score-transformed mean decrease in accuracy.

>>>>

**Figure 4: Relative importance of taxa with significant ORs in Random Forest models for differentiating between individuals with normal colons and carcinomas using stool samples (A) or unmatched tissue samples (B).** The colors indicate the z-transformed (i.e. mean of 0.0 and standard deviation of 1.0) mean decrease in accuracy values calculated from the model for each study. The taxa are ranked by their mean z-score-transformed mean decrease in accuracy.

**Figure 5: Comparison of Random Forest modeling approaches to classify individuals as having normal colons or adenomas (A) or carcinomas (B) when training the models using the taxa with significant ORs, all taxa in a community, or all OTUs in a community when using stool samples.** No taxa had a significant OR associated with the presence of adenomas using stool samples. The black line represents the median AUC for the respective group. The dashed gray line indicates an AUC of 0.5.

### 2018-03-15 submission/manuscript.tex
change fig legend order

@@

\textbf{Table 1: Total Individuals in each Study Included in the Fecal
Analysis}

@@

\textbf{Table 2: Studies with Tissue Samples Included in the Analysis}

@@

\textbf{Figure 1: Significant Bacterial Community Metrics for Adenoma or
Carcinoma in Feces} A) Adenoma evenness. B) Carcinoma evenness. C)
Carcinoma Shannon diversity. Blue represents controls and red represents
either adenoma (panel A) or carcinoma (panel B and C). The black lines
represent the median value for each repsective group.

\textbf{Figure 2: Odds Ratio for Adenoma or Carcinoma based on Bacterial
Community Metrics in Feces.} A) Community-based odds ratio for adenoma.
B) Community-based odds ratio for carcinoma. Colors represent the
different variable regions used within the respective study.

\textbf{Figure 3: The AUC of Indivdiual Significant OR Taxa to classify
Carcinoma.} A) Fecal samples. B) Unmatched tissue samples. The larger
circle represents the median AUC of all studies and the smaller circles
represent the individual AUC for a particular study. The dotted line
denotes an AUC of 0.5.

\textbf{Figure 4: Most Important Members in Significant OR Taxa
Carcinoma Models.} A) Common taxa in the top 10 percent for carcinoma
Random Forest fecal-based models. B) Common taxa in the top 10 percent
for carcinoma Random Forest unmatched tissue-based models. Blue
represents less important and red represents more important to the
Random Forest Model. White means that particular taxa was not in the top
10\%.

\textbf{Figure 5: Fecal Sample Random Forest Model Train AUCs.} A)
Adenoma random forest model AUCs between all genera, all OTU, and select
model based on significant OR taxa. B) Carcinoma random forest model
AUCs between all genera, all OTU, and select model based on significant
OR taxa. The black line represents the median AUC for the respective
group. If no values are present in the singificant OR taxa group then
there were no significant taxa identified and no model was tested.

\textbf{Figure 6: Fecal Sample Random Forest Genus-Based Model Test
AUCs.} A) Test AUCs of adenoma models using all genera across study. B)
Test AUCs of carcinoma models using all genera or significant OR taxa
only. The black line represents the AUC at 0.5. The red lines represent
the median AUC of all test AUCs for a specific study.

@@

\textbf{Figure S1: Odds Ratio for Adenoma or Carcinoma based on
Bacterial Community Metrics in Tissue.} A) Community-based odds ratio
for adenoma. B) Community-based odds ratio for carcinoma. Colors
represent the different variable regions used within the respective
study.

\textbf{Figure S2: Tissue Random Forest Model Train AUCs.} A) Adenoma
random forest model AUCs between all genera, all OTU, and select model
based on significant OR taxa in unmatched and matched tissue. B)
Carcinoma random forest model AUCs between all genera, all OTU, and
select model based on significant OR taxa in unmatched and matched
tissue. The black line represents the median AUC for the respective
group. If no values are present in the singificant OR taxa group then
there were no significant taxa identified and no model was tested.

\textbf{Figure S3: Most Common Taxa Across Carcinoma Full Community
Fecal Sample Study Models.} A) Common taxa in the top 10 percent for
carcinoma Random Forest all taxa-based models. B) Common taxa in the top
10 percent for carcinoma Random Forest all OTU-based models. Blue
represents less important and red represents more important to the
Random Forest Model. White means that particular taxa was not in the top
10\%.

\textbf{Figure S4: Most Common Genera Across Full Community Tissue Study
Models.} A) Common genera in the top 10 percent for matched carcinoma
Random Forest all genera-based models. B) Common genera in the top 10
percent for unmatched carcinoma Random Forest all genera-based models.
C) Common genera in the top 10 percent for matched carcinoma Random
Forest all OTU-based models. D) Common genera in the top 10 percent for
unmatched carcinoma Random Forest all OTU-based models. Blue represents
less important and red represents more important to the Random Forest
Model. White means that particular taxa was not in the top 10\%.

\textbf{Figure S5: Tissue Random Forest Genus-Based Model Test AUCs.} A)
Test AUCs of adenoma models using all genera across study. B) Test AUCs
of carcinoma models using all genera for matched tissue studies. C) Test
AUCs of carcinoma models using all genera or significant OR taxa only
for unmatched tissue studies The black line represents the AUC at 0.5.
The red lines represent the median AUC of all test AUCs for a specific
study.

>>>>

\usepackage{lscape}
\newcommand{\blandscape}{\begin{landscape}}
\newcommand{\elandscape}{\end{landscape}}

@@

\textbf{Table 1: Characteristics of the datasets included in the
fecal-based analysis}

@@

\textbf{Table 2: Characteristics of the datasets included in the
tissue-based analyses}

@@

\textbf{Figure 1: Comparison of alpha diversity indices that were
significant between individuals with normal colons, and those with
adenomas or carcinomas using data collected from fecal samples} A)
Comparison of evenness between individuals with normal colons and
adenomas. B) Comparison of evenness between individuals with normal
colons and carcinomas. C) Comparison of Shannon diversity between
individuals with normal colons and carcinomas. Blue points represent
individuals with normal colons and red points represent individuals with
either adenomas (panel A) or carcinomas (panel B and C). The black lines
represent the median value for each group.

\textbf{Figure 2: Comparison of odds ratios calculated using alpha
diversity community metrics associated with the presence of adenomas (A)
or carcinoma (B) relative to those in individuals with normal colons
using data collected from stool samples.}

\textbf{Figure 3: AUC values when classifing individuals as having
normal colons or carcinomas using taxa with significant ORs when using
stool samples (A) and unmatched tissue samples (B).} We did not identify
any taxa as having a significant OR to differentiate individuals with
normal colons and adenomas or using matched tissue samples. The large
black circles represent the median AUC of all studies and the smaller
circles represent the individual AUC for a particular study. The dotted
line denotes an AUC of 0.5.

\textbf{Figure 4: Relative importance of taxa with significant ORs in
Random Forest models for differentiating between individuals with normal
colons and carcinomas using stool samples (A) or unmatched tissue
samples (B).} The colors indicate the z-transformed (i.e.~mean of 0.0
and standard deviation of 1.0) mean decrease in accuracy values
calculated from the model for each study. The taxa are ranked by their
mean z-score-transformed mean decrease in accuracy.

\textbf{Figure 5: Comparison of Random Forest modeling approaches to
classify individuals as having normal colons or adenomas (A) or
carcinomas (B) when training the models using the taxa with significant
ORs, all taxa in a community, or all OTUs in a community when using
stool samples.} No taxa had a significant OR associated with the
presence of adenomas using stool samples. The black line represents the
median AUC for the respective group. The dashed gray line indicates an
AUC of 0.5.

\textbf{Figure 6: Testing of Random Forest models to classify
individuals as having normal colons or adenomas (A) or carcinomas (B)
when using sequence data obtained from stool samples.} Models were
trained on data from each study (Figure 5) and tested on the other
studies. The black lines represent the median AUC of all test AUCs for a
specific study. The dashed gray line represents the AUC at 0.5.

@@

\textbf{Figure S1: Comparison of Odds Ratios associated with normal
colons or adenomas (A) or carcinomas (B) calculated using alpha
diversity indices with sequence data generated from tissue samples.} The
pooled results are from the aggregation of data across all studies. The
horizontal lines indicate the 95\% confidence interval for the OR.

\textbf{Figure S2: Relative importance of taxa (A) and OTUs (B) in
Random Forest models for differentiating between individuals with normal
colons and carcinomas using stool samples.} These taxa and OTUs were
among the top 10\% most important features in each model. The colors
indicate the z-transformed (i.e.~mean of 0.0 and standard deviation of
1.0) mean decrease in accuracy values calculated from the model for each
study. The taxa are ranked by their mean z-score-transformed mean
decrease in accuracy.

\textbf{Figure S3: Relative importance of taxa (A, B) and OTUs (C, D) in
Random Forest models for differentiating between individuals with normal
colons and carcinomas using matched (A, C) and unmatched (B, D) tissue
samples.} hese taxa and OTUs were among the top 10\% most important
features in each model. The colors indicate the z-transformed (i.e.~mean
of 0.0 and standard deviation of 1.0) mean decrease in accuracy values
calculated from the model for each study. The taxa are ranked by their
mean z-score-transformed mean decrease in accuracy.

\textbf{Figure S4: Comparison of Random Forest modeling approaches to
classify individuals as having normal colons or adenomas (A) or
carcinomas (B) when training the models using the taxa with significant
ORs, all taxa in a community, or all OTUs in a community when using data
from tissue samples.} No taxa had a significant OR associated with the
presence of adenomas using tissue samples. The black line represents the
median AUC for the respective group. The dashed gray line indicates an
AUC of 0.5.

\textbf{Figure S5: Testing of Random Forest models to classify
individuals as having normal colons or adenomas (A) or carcinomas (B, C)
when using sequence data obtained from tissue samples.} Models were
trained on data from each study (Figure S5) and tested on the other
studies. The black lines represent the median AUC of all test AUCs for a
specific study. The dashed gray line represents the AUC at 0.5.
